---
title: "Introduction to nepalidate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to nepalidate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(nepalidate)
```

## Overview

The `nepalidate` package converts dates between the Nepali
Bikram Sambat (BS) calendar and the Gregorian (AD) calendar.
It supports BS years 1970--2099 (AD 1913--2043).

## Basic Conversion

Convert BS to AD with `bs_to_ad()` --- returns a `Date` object:

```{r}
bs_to_ad("2080-05-15")
```

Convert AD to BS with `ad_to_bs()` --- returns a `bs_date` object:

```{r}
ad_to_bs("2023-09-01")
```

Both accept `Date` objects and `POSIXct`:

```{r}
ad_to_bs(Sys.Date())
```

## Working with Vectors

Both functions are fully vectorized and handle `NA` values:

```{r}
ad_dates <- c("2023-01-01", "2023-06-15", NA, "2023-12-31")
ad_to_bs(ad_dates)
```

## The bs_date Class

`ad_to_bs()` returns a `bs_date` object with useful methods.

### Formatting

```{r}
d <- ad_to_bs("2023-09-01")
format(d, "%d %B %Y")
```

### Arithmetic

Add or subtract days:

```{r}
d <- bs_date("2080-05-15")
d + 30
d - 10
```

Compute the difference between two BS dates (in days):

```{r}
bs_date("2080-06-15") - bs_date("2080-05-15")
```

### Comparison

```{r}
bs_date("2080-06-01") > bs_date("2080-05-15")
```
### Convert to AD Date

```{r}
as.Date(bs_date("2080-05-15"))
```

## Working with Data Frames

```{r}
df <- data.frame(
  id = 1:3,
  reg_date = as.Date(c("2023-01-15", "2023-06-20", "2023-11-10"))
)

df$bs_date <- ad_to_bs(df$reg_date)
df
```

## Month Names

```{r}
get_nepali_month_name(c(1, 5, 12))
get_nepali_month_name(1, "ne")
```
